### 2022-05-06 Markdown/Manuscript/Child_RMDs/Doodles.Rmd
still on intro

>>>>

CLIPPINGS
It is frequently claimed that better approaches to the organising and financing of mental health service systems can generate significant gains in population health (Saxena, 2007; Chisholm, 2016).  Over the last two decades, policymakers in Australia have acted on these claims by investing in a series of major systemic mental health reforms, with improving the mental health of young people a major focus (McGorry, 2017). The scope and complexity of these reforms has increased over time, with the Victorian Government’s current plan to overhaul mental health service delivery (Victoria, 2021) arguably the most ambitious to-date.

 Although Australian mental health policy and system design processes has a formally specified role for a robust model-based decision aid, this tool,  the National Mental Health Services Planning Framework, has limited capacity to meet many of the core decision-support requirements of systems planners (Commission, 2020). These requirements are diverse and extensive, but are likely to include models that can:

•	anticipate the distribution and evolution of risk factors and need for mental health care by region and population group;

•	identify the budget, equity and efficiency impacts of alternative resource allocations between mental health programs;

•	provide guidance about how services can be optimally configured to achieve outcomes such as minimising wait times, maximising health gain, meeting patient preferences and improving societal welfare; and

•	predict how the supply, location and behaviour of mental healthcare providers change in response to alternative incentives and conditions.

Although a number of these or similar types of models have been developed for the Australian context, the validity, utility and accessibility of these models vary considerably.

and to anticipate and mitigate risks to the successful implementation of these reforms

Developing theories that produce valid and useful predictions about the potential impacts and feasibility of phased, multi-component programs of policy and service system reform is most likely a major and long term challenge. This undertaking might become more tractable if it could be chunked into smaller tasks and progressed by a research community that, even when working independently, routinely test, re-use and extend each other's work.

However, developing  This leads to computational model issue. This leads to open source.

The absence of common standards and taxonomies make it hard to develop these more extensive models by simply combining existing mental health models.

 which can be used to generate credible predictions about how these systems might respond to phased, multi-component redesign strategies

Common standards and frameworks for how computational models are developed could address some of these challenges [@RN73].

Outputs from discrete projects should be reproducible, generalisable and able to be linked like Lego bricks to progressively assemble more complex and realistic representations of the systems being modelled.

 Similarly,

These knowledge gaps may partly explain why debates within the mental health policymaking community extend beyond normative issues (what should policy aim to achieve) to technical questions (which system financing and organising choices are most likely to achieve specified objectives). A more mature mental health systems modelling field might see the technical aspects of re-engineering mental health systems subject to the same stringent scientific standards as those required for the development of therapeutic goods. Yet,

If these issues can be successfully addressed, there are potential benefits to both decision makers (who will have access to more relevant and trustworthy model-based decision aids) and modellers (for whom complex models will become more tractable through improved access to relevant code and data).

Access to some model-based decision aids and their constituent algorithms and data is highly restricted. The relevance of models to current and future decision contexts can also be limited as mental health models are rarely routinely updated with new data and improved algorithms.

In addition to issues relating to improving the quality and usefulness of models that address narrowly defined topics, there are additional challenges concerning the development of models with much broader scope.

A number of high income countries have begun to increase public expenditures on mental health services[@RN26], and in some cases have embraced ambitious programs of major structural reforms to mental health service systems[REFSGOHERE]. However

### 2022-05-06 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
still on intro

---
title: "MS_Intro_Short"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`

>>>>

---
title: "MS_Intro_Short"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Mental disorders impose high health, social and economic burdens worldwide [@RN8][@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equiped for this challenge[@RN22][@RN23]. The large and widespread additional mental health burdens recently observed during the COVID pandemic[@20211700] and predicted as a potential future consequence of global heating [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorder[@RN11].

Realising significant improvements in population mental health may in part depend on gaining better understanding of the systems in which mental disorder emerges and is treated [@Fried2020]. Currently, the theoretical basis for understanding these systems is weak [@RN2111]. Strikingly, it remains unclear why increased investments in mental health care have yet to discernably reduce the prevalence and burden of mental disorders[@RN26].  The literature about how the requirements, characteristics and performance of mental health services are shaped by spatiotemporal context is underdeveloped [@RN42]. There is insufficient evidence to identify the social determinants of mental disorders most amenable to preventative interventions, and for which population sub-groups such interventions would be most effective [@RN43].

Mental health simulation studies rarely explore the features and behaviours of complex systems [@RN73], with mental health economic models predominantly addressesing issues relating to the affordability and value for money of individual programs [@RN34]. Single purpose models that assume static systems, even when sufficiently robust to be formally incorported into Government resource planning processes, may be inadequate for the decision support needs of policymakers and service planners seeking to successfully identify, prioritise, sequence and target multiple elements of complex reform programs [@PC2020].

Multi-application, dynamic systems modelling approaches can provide insights about inter-dependencies between candidate policy settings that static models of isolated scenarios cannot [@Occhipinti2021]. However, greater use of these types of models may require adaptation on the part of funders, modellers and decision-makers. The complexity of multi-application systems models may make them particularly prone to errors [@Saltelli2019], thus requiring greater investments in model transparency and validation [@Eddy2012]. These types of models can be highly resource intensive to develop and may also never be truly "finished", instead requiring ongoing updates to remain relevant to evolving decision contexts [@Jenkins2021] and to meet additional feature requests from end-users. The development, validation and maintenance of these types of models may be simply too onerous and long term a burden to remain the responsibility of a single modelling team and might benefit from collaborations across multiple modelling teams, and perhaps the general mental health modelling field. Similarly, more attention to developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure multi-purpose models can address priority decision topics and have user-interfaces that meet the needs of non-technical users.

Open source frameworks, which have been recommended for the development of mental health modelling field (Long, 2018), have the potential to help address a number of these challenges. Open source approaches can promote common standards and interoperability, improve model transparency and facilitate routine updates to rectify errors, update data and add features. They may also have educational value in developing the health systems modelling field (Dunlop, 2017).

However, open source computational models also present a range of conceptual, technical, legal and resourcing issues compared to more traditional approaches to model development.

However, of the three components that comprise a mental health computer simulation model - the conceptual model, the mathematical model and the computational model - only the first two are typically developed and reported in a manner that facilitates reuse, refinement and extension by others. This is problematic as the code implementation is where the theory represented by a simulation model is most fully elaborated. In the limited cases where code is made publicly available it is typically in the form of programs (an algorithm that performs one task) rather than code libraries (where the primary algorithm is decomponsed into multiple component algorithms, that can be re-used for a potentially large number of other purposes).

```{r echo=F}
# ISPOR T & V: 
# a multiapplication model, transparency, validation,
# For a multiapplication model, transparency, validation,
# and reporting are ongoing processes. The multiapplica-
# tion model is described (transparency)24 and validated,25
# and the descriptions and validations are continuously
# updated as science and the model evolve.26 In addition,
# each instantiation of the model is described, validated,
# and reported as each application is done.27
```

### 2022-05-06 Markdown/Manuscript/Parent_PDF/Parent_PDF.Rmd
still on intro

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_book: 
    base_format: rticles::arxiv_article
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
  bookdown::pdf_document2: 
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
    toc: no
## CUSTOM_AUTHORS_PLACEHOLDER ##
params:
  X: NULL
abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---

>>>>

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_book: 
    base_format: rticles::arxiv_article
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
  bookdown::pdf_document2: 
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
    toc: no
## CUSTOM_AUTHORS_PLACEHOLDER ##
params:
  X: NULL
abstract: "There is strong in principle support for open source health economic models, but practical barriers limit their availability. We propose a set of principles and standards for the implementation of open source health economic models that are TIMELY - Transparent, Iterative, Modular, Epitomised and Yielding. We then describe a software framework that we have developed for developing TIMELY models in youth mental health and illustrate this framework with an open source utility mapping project."
# abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---
